% ************************** Thesis Acknowledgements **************************
{
\cleardoublepage
\setsinglecolumn
\chapter*{\centering \LARGE Acknowledgements}
\thispagestyle{empty}

A doctoral thesis in the life sciences might have a single author, but almost always has a great many direct and indirect contributors. Mine is no exception, and so I have a number of people to thank.

First, my sincere thanks to Dario Valenzano, who welcomed me into his research group in 2014 and has always been an enthusiastic, attentive and understanding supervisor. None of the work in this thesis would have been possible without him, and I would be a far poorer scientist without his mentorship and support.

Many members of the Valenzano group have actively contributed to this thesis. Patrick Smith performed the microbiota-transfer studies that were the inspiration for the project in the first place and supervised and mentored me through my first and most difficult year. Joanna Dodzian generously bred and provided the fish cohorts I used for the pilot and ageing experiments in \Cref{chap:igseq}. Rongfeng (Ray) Cui performed the genome assembly that laid the foundation for the work in \Cref{chap:locus} and provided a great deal of advice on both experimental procedures and computational analysis. My student supervisees, Linda Zirden, Lena Schlautmann, Pascha Hokama, and Davina Patel all made essential contributions to developing and testing the turquoise-killifish immunoglobulin-sequencing protocol; you were all great. 

Our tireless and wonderful technician, Aleksandra Placzek, has been an essential part of the \igseq project in the lab and performed several of the library preps; many of my favourite parts of this thesis would not have been possible without her. Most recently, Michael Poeschla has joined the repertoire-sequencing team and helped prepare the libraries for the final \igseq experiment in \Cref{chap:igseq}; I look forward to seeing where he takes the project in the future. In additional to Dario and Michael, my fellow students David Willemsen, Jens Seidel, Miriam Popkes and Daniel Davila all helped proofread the manuscript for this thesis; Miriam and especially Jens also did the lion's share of the work translating my abstract into German. I am sincerely grateful to all of them, and to all other past and present members of the Valenzano lab, for their help, support, and enthusiasm, and for making the group such a pleasant and friendly place in which to work.

A number of external collaborators and other scholars also made vital contributions to this work. Kathrin Reichwald provided me with essential BAC clones and a substantial amount of important advice. John Beausang, Aleksandra Walczak, Thierry Mora and Susana Magadan helped me get started in the enormous and somewhat intimidating world of immune-repertoire analysis, and Aleksandra in particular has repeatedly provided invaluable advice, introductions and insight. Jason Vander Heiden has tirelessly answered my many questions about the Immcantation pipeline and even made some changes to it at my request, while Quentin Marcou has been an invaluable help in getting to grips with his fantastic software \program{IGoR}. Tsimafei Padvitski and Davi Silva generously provided additional proofreading. Many other researchers have responded to my pestering emails, given advice where I needed it, and generally been friendly, welcoming and generous with their time. I am grateful to you all.

The co-ordination team of the Cologne Graduate School of Ageing Research has consistently been amazing, and I have repeatedly been impressed at the depth of their dedication, sympathy and willingness to help us students with all our big and little problems. My sincere thanks to Daniela Morick, Doris Birker and Jenny Ostermann for all their help and support. Thanks also to Ruth Willmott, for her invaluable soft-skills and careers coaching; Jorge Boucas, Franzi Metge, Daniel Rosskopp and the rest of the bioinformatics core facility for their patient help and advice; and all the other people I have unaccountably failed to mention here whose assistence, friendship and support I have relied on throughout this lengthly process.

My family, of course, have been a constant source of love and support. Mum, Dad, Anna, thank you for everything. I love you. My love and thanks also to the rest of my family, especially Ron and Julie, for their enduring support and belief in me and my research. 

Finally, my deepest and sincerest thanks to Simon Woolf, without whom this entire long adventure would have been a great deal more difficult and a great deal less fun.
}
